---
title: "Analysis Business Research"
author: "Melle Stephen Oudshoorn"
date: "`r Sys.Date()`"
output: html_document
---

## Loading libraries

```{r loading packages message = FALSE, warning = FALSE}
library(dplyr)
library(readxl)
library(tidyr)
library(readr)
library(ggplot2)
library(ltm)
library(janitor)
library(car)
library(gridExtra)
library(grid)
library(lavaan)
```

## Preparations

```{r load and mutate data}
# Read in the data
data <- read_excel("Master File.xlsx")

# Clean column names to remove special characters
data <- janitor::clean_names(data)  # janitor::clean_names() makes all names lowercase, replaces spaces with underscores, etc.

# Define selected columns based on the cleaned names
selected_columns <- c("in_what_kind_of_place_do_you_typically_work",
                      "in_your_main_job_have_you_spent_at_least_one_full_day_in_the_last_four_weeks_working",
                      "how_frequently_did_you_work_from_home_in_the_last_four_weeks",
                      "i_currently_have_a_good_balance_between_the_time_i_spend_at_work_and_the_time_i_have_available_for_non_work_activities",
                      "i_have_sufficient_time_away_from_my_job_to_maintain_adequate_well_being",
                      "i_am_satisfied_with_my_work_life_balance",
                      "my_personal_life_suffers_because_of_work_43",
                      "when_i_get_home_i_am_often_too_exhausted_to_participate_in_family_activities",
                      "the_problem_solving_behaviours_i_use_in_my_job_are_not_effective_in_resolving_problems_at_home",
                      "the_time_i_spend_on_family_responsibilities_often_interferes_with_my_work",
                      "due_to_stress_at_home_i_am_often_preoccupied_with_family_matters_at_work",
                      "the_behaviours_that_work_for_me_at_home_do_not_seem_to_be_effective_at_work",
                      "what_is_your_gender",
                      "what_is_your_age",
                      "do_you_work_right_now_or_have_you_worked_in_the_last_year",
                      "in_what_job_industry_was_is_this_work",
                      "how_many_years_of_experience_do_you_have_in_your_current_last_years_job_numbers_only",
                      "what_is_the_highest_level_of_education_that_you_have_completed",
                      "what_is_your_type_of_employment")

# Select the columns of interest
data <- data %>% 
  dplyr::select(all_of(selected_columns))

# Rename the selected columns
data <- data %>%
  rename(location = in_what_kind_of_place_do_you_typically_work,
         work_last_4weeks = in_your_main_job_have_you_spent_at_least_one_full_day_in_the_last_four_weeks_working,
         home_work_freq = how_frequently_did_you_work_from_home_in_the_last_four_weeks,
         WLB1 = i_currently_have_a_good_balance_between_the_time_i_spend_at_work_and_the_time_i_have_available_for_non_work_activities,
         WLB2 = i_have_sufficient_time_away_from_my_job_to_maintain_adequate_well_being,
         WLB3 = i_am_satisfied_with_my_work_life_balance,
         WFC1 = my_personal_life_suffers_because_of_work_43,
         WFC2 = when_i_get_home_i_am_often_too_exhausted_to_participate_in_family_activities,
         WFC3 = the_problem_solving_behaviours_i_use_in_my_job_are_not_effective_in_resolving_problems_at_home,
         FWC1 = the_time_i_spend_on_family_responsibilities_often_interferes_with_my_work,
         FWC2 = due_to_stress_at_home_i_am_often_preoccupied_with_family_matters_at_work,
         FWC3 = the_behaviours_that_work_for_me_at_home_do_not_seem_to_be_effective_at_work,
         gender = what_is_your_gender,
         age = what_is_your_age,
         working_status = do_you_work_right_now_or_have_you_worked_in_the_last_year,
         industry = in_what_job_industry_was_is_this_work,
         experience = how_many_years_of_experience_do_you_have_in_your_current_last_years_job_numbers_only,
         education_level = what_is_the_highest_level_of_education_that_you_have_completed,
         employment_type = what_is_your_type_of_employment)

# Check the renamed data
head(data)
```

```{r create working at home variable}
determine_work_type <- function(location, work_last_4weeks, home_work_freq) {
  if (location == 1) {
    return(1)  # Works from Home
  } else if (location %in% 2:8 && work_last_4weeks %in% 1:2 && home_work_freq == 1) {
    return(2)  # Works Hybrid
  } else if (location %in% 2:8 && work_last_4weeks == 3) {
    return(3)  # Works on location
  } else if (location %in% 2:8 && work_last_4weeks %in% 1:2 && home_work_freq == 2) {
    return(3)  # Works on location
  } else {
    return(3)
  }
}

data <- data %>%
  mutate(RW = mapply(determine_work_type, location, work_last_4weeks, home_work_freq)) %>% 
  dplyr::select(-c("location", "work_last_4weeks", "home_work_freq")) %>% 
  na.omit() # removed 4 observations! check if this can be prevented.

data <- data %>%
  mutate(
    RW = 4 - RW,
    WLB1 = 6 - WLB1,
    WLB2 = 6 - WLB2,
    WLB3 = 6 - WLB3,
    WFC1 = 6 - WFC1,
    WFC2 = 6 - WFC2,
    WFC3 = 6 - WFC3,
    FWC1 = 6 - FWC1,
    FWC2 = 6 - FWC2,
    FWC3 = 6 - FWC3
  )
```

```{r converting data types}
library(dplyr)

# Convert categorical variables to factors with appropriate levels
data <- data %>%
  mutate(
    gender = factor(gender, levels = c(1, 2, 3), labels = c("Male", "Female", "Other")),
    education_level = factor(education_level, levels = c(1, 2, 3, 4, 5), 
                             labels = c("Primary", "Secondary", "Bachelor's", "Master's", "Doctorate")),
    employment_type = factor(employment_type, levels = c(1, 2, 3, 4, 5, 6), 
                             labels = c("Full-time", "Part-time", "Internship", "Temporary", "Freelance", "Other"))
  )

```

```{r crohnbach alpha}

#initialise list of multi-item constructs
constructs <- list(
  WLB = c("WLB1", "WLB2", "WLB3"),
  WFC = c("WFC1", "WFC2", "WFC3"),   
  FWC = c("FWC1", "FWC2", "FWC3")
)

#create function to calculate the alpha
calculate_cronbach_alpha <- function(items) {
  alpha_result <- cronbach.alpha(data[, items], na.rm = TRUE)
  return(alpha_result)
}

#loop
cronbach_results <- lapply(constructs, calculate_cronbach_alpha)

for (construct_name in names(cronbach_results)) {
  cat("Cronbach's Alpha for", construct_name, ":", cronbach_results[[construct_name]]$alpha, "\n")
}
```

The Alpha for WLB and FWC are acceptable, unfortunately WFC does not meet the requirement of at least 0.7. However, when I remove any question, the alpha just gets lowered. Because of that I will accept this bias, and move on with all of these items for WFC.

```{r create the scores}
data <- mutate(data, WLB_score = rowMeans(data[, c("WLB1", "WLB2", "WLB3")], na.rm = TRUE))
data <- mutate(data, WFC_score = rowMeans(data[, c("WFC1", "WFC2", "WFC3")], na.rm = TRUE))
data <- mutate(data, FWC_score = rowMeans(data[, c("FWC1", "FWC2", "FWC3")], na.rm = TRUE))
data <- mutate(data, Conflict_score = rowMeans(data[, c("WFC1", "WFC2", "WFC3", "FWC1", "FWC2", "FWC3")], na.rm = TRUE))
```

## Description 

### Demographics

```{r gender}
print((table(data$gender) / sum(table(data$gender))) * 100)
```

```{r age}
summary(data$age)
hist.age <- ggplot(data = data, mapping = aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black") + # Histogram 
  labs(title = "Age Distribution",
       x = "Age",
       y = "Count") + # Labels for axes
  theme_minimal() # A minimal theme for a cleaner look
sd(data$age)
```

```{r experience}
summary(data$experience)
hist.exp <- ggplot(data = data, mapping = aes(x = experience)) +
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") + # Histogram
  labs(title = "Experience Distribution",
       x = "Years",
       y = "Count") + # Labels for axes
  theme_minimal() # A minimal theme for a cleaner look
```

```{r}
grid.arrange(hist.age, hist.exp, ncol = 2)
```

```{r education level}
# Ensure education_level is a factor
data$education_level <- as.factor(data$education_level)

# Plot the education level distribution
ggplot(data = data, mapping = aes(x = education_level)) +
  geom_bar(fill = "lightgreen", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(title = "Education Level Distribution",
       x = "Education Level",
       y = "Count") +
  theme_minimal() +
  scale_x_discrete(labels = c("1 Primary", 
                               "2 Secondary", 
                               "3 Bachelor's", 
                               "4 Master's", 
                               "5 Doctorate")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))

# Count and calculate percentages
data %>%
  count(education_level) %>%
  mutate(percentage = (n / sum(n)) * 100)

```

```{r employment type}
# Plot the employment type distribution
ggplot(data = data, mapping = aes(x = employment_type)) +
  geom_bar(fill = "lightgreen", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(title = "Employment Type",
       x = "Type",
       y = "Count") +
  theme_minimal() +
  scale_x_discrete(labels = c("1 Full-time", 
                               "2 Part-time", 
                               "3 Internship", 
                               "4 Temporary", 
                               "5 Freelance",
                               "6 Other")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))

# Count and calculate percentages
data %>%
  count(employment_type) %>%
  mutate(percentage = (n / sum(n)) * 100)
```

### Variables of interest

```{r summary statistics}
print("Work Life Balance")
summary(data$WLB_score)
sd(data$WLB_score)
print("Work Family Conflict score")
summary(data$WFC_score)
sd(data$WFC_score)
print("Family Work Conflict score")
summary(data$FWC_score)
sd(data$FWC_score)
```
```{r}
ggplot(data) +
  # Creating individual box plots for each score variable
  geom_boxplot(aes(x = "WLB Score", y = WLB_score, fill = "WLB Score"), width = 0.4) +
  geom_boxplot(aes(x = "WFC Score", y = WFC_score, fill = "WFC Score"), width = 0.4) +
  geom_boxplot(aes(x = "FWC Score", y = FWC_score, fill = "FWC Score"), width = 0.4) +
  
  # Setting up titles and labels
  labs(title = "Summary Statistics of Work and Family Scores", x = "Category", y = "Score") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("WLB Score" = "lightblue", "WFC Score" = "lightgreen", "FWC Score" = "lightcoral"))
```


```{r remote work}
ggplot(data = data, mapping = aes(x = factor(RW,
                                              levels = c(1,2,3), 
                                              labels = c("1 On location", 
                                                         "2 Hybrid", 
                                                         "3 Remotely" 
                                                        )))) +
  geom_bar(fill = "lightgreen", color = "black") +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(title = "Work location",
       x = "Type",
       y = "Count") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
```

```{r}
WLB_density <- ggplot(data, aes(x = WLB_score)) +
  geom_density(bins = 10, fill = "lightgreen", color = "black") +
  labs(title = "Work-Life Balance", x = "WLB Score") +
  theme_minimal() +
  theme(axis.title.y = element_text(size = 12),
        axis.text.y = element_blank(),          
        axis.ticks.y = element_blank())         

WFC_density <- ggplot(data, aes(x = WFC_score)) +
  geom_density(bins = 10, fill = "lightblue", color = "black") +
  labs(title = "Work-Family Conflict", x = "WFC Score") +
  theme_minimal() +
  theme(axis.title.y = element_blank(),       
        axis.text.y = element_blank(),        
        axis.ticks.y = element_blank())         

FWC_density <- ggplot(data, aes(x = FWC_score)) +
  geom_density(bins = 10, fill = "lightyellow", color = "black") +
  labs(title = "Family-Work Conflict", x = "FWC Score") +
  theme_minimal() +
  theme(axis.title.y = element_blank(),         
        axis.text.y = element_blank(),          
        axis.ticks.y = element_blank())      

grid.arrange(WLB_density, WFC_density, FWC_density, ncol = 3, 
             top = textGrob("Density plots", gp = gpar(fontsize = 14)), 
             bottom = textGrob("", gp = gpar(fontsize = 12)))
```

## Correlation analysis

```{r}
correlation_matrix <- cor(data[c("WLB_score", "WFC_score", "FWC_score", "RW")], use = "complete.obs")
correlation_matrix
```

## Exploratory Factor Analysis (EFA)

```{r setup EFA}
# Load the psych library for EFA functions
library(psych)

# Select items for factor analysis (in this example, using all WLB, WFC, and FWC items)
efa_items <- data %>% dplyr::select(WLB1, WLB2, WLB3, WFC1, WFC2, WFC3, FWC1, FWC2, FWC3)

# Check the adequacy of your data for EFA (Kaiser-Meyer-Olkin Measure of Sampling Adequacy)
KMO_result <- KMO(efa_items)
cat("Kaiser-Meyer-Olkin Measure of Sampling Adequacy:", KMO_result$MSA, "\n")
print(KMO_result)

# Bartlett's Test of Sphericity
bartlett_result <- cortest.bartlett(cor(efa_items), n = nrow(data))
cat("Bartlett's Test of Sphericity:\n")
print(bartlett_result)
```

Chi square is very high which means that the data is a very good fit for Factor analysis.

```{r}
# Scree plot to help determine the number of factors
fa.parallel(efa_items, fm = "ml", fa = "fa")

# Conduct Exploratory Factor Analysis
# Here we assume 2 factors based on Scree Plot or Kaiser’s criterion (adjust the number if necessary)
efa_result <- fa(efa_items, nfactors = 3, rotate = "varimax", fm = "ml")

# Display factor loadings
print(efa_result$loadings, cutoff = 0.3)

```

Scree plot suggests that I should extract 2 components

Factor 1 (ML1) explains 28.3% of the variance.
Factor 2 (ML2) explains 24.3% of the variance.
Cumulative Variance: Together, these factors explain 52.6% of the variance, which is generally acceptable for social science constructs

```{r}
# Specify CFA model based on EFA results
cfa_model <- '
  WLB =~ WLB1 + WLB2 + WLB3
  FWC =~ WFC1 + WFC2 + WFC3 + FWC1 + FWC2 + FWC3
'

# Fit the model
fit <- cfa(cfa_model, data = data)

# Review model fit
summary(fit, fit.measures = TRUE, standardized = TRUE)

```

Extremely low p-value so highly statistically significant. WFC3 has a low loading (0.363), which means it might not be strongly related to the latent factor FWC.

## Hypothesis testing

### Linear models

```{r H-1: Remote working negatively affects work-life balance by blurring boundaries between work and personal life}
# Perform a linear regression for H-1
model_H1 <- lm("WLB_score ~ RW + experience + gender + education_level + employment_type", data = data)

summary(model_H1)

# Boxplot to visualize the relationship between remote work type and work-life balance
ggplot(data, aes(x = as.factor(RW), y = WLB_score)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Effect of Remote Working on Work-Life Balance",
       x = "Work Type (1 = Remote, 2 = Hybrid, 3 = On Location)",
       y = "Work-Life Balance Score") +
  theme_minimal()

```

```{r H-2.1: Remote working will influence work-family conflict negatively}
# Perform a linear regression for H-2.1
model_H21 <- lm("WFC_score ~ RW + gender + education_level + employment_type", data = data)

summary(model_H21)

# Boxplot to visualize the relationship between remote work type and work-family conflict
ggplot(data, aes(x = as.factor(RW), y = WFC_score)) +
  geom_boxplot(fill = "lightcoral") +
  labs(title = "Effect of Remote Working on Work-Family Conflict",
       x = "Work Type (1 = Remote, 2 = Hybrid, 3 = On Location)",
       y = "Work-Family Conflict Score") +
  theme_minimal()

```

```{r H-2.2: Remote working will influence family-work conflict negatively}
# Perform a linear regression for H-2.2
model_H22 <- lm("FWC_score ~ RW + gender + education_level + employment_type", data = data)

summary(model_H22)

# Boxplot to visualize the relationship between remote work type and work-family conflict
ggplot(data, aes(x = as.factor(RW), y = FWC_score)) +
  geom_boxplot(fill = "lightcoral") +
  labs(title = "Effect of Remote Working on Work-Family Conflict",
       x = "Work Type (1 = Remote, 2 = Hybrid, 3 = On Location)",
       y = "Work-Family Conflict Score") +
  theme_minimal()
```

```{r H-2.3 combined conflicts}
# Perform a linear regression for H-2.3
model_H23 <- lm("Conflict_score ~ RW + gender + education_level + employment_type", data = data)

summary(model_H23)

# Boxplot to visualize the relationship between remote work type and work-family conflict
ggplot(data, aes(x = as.factor(RW), y = Conflict_score)) +
  geom_boxplot(fill = "lightcoral") +
  labs(title = "Effect of Remote Working on Work-Family Conflict",
       x = "Work Type (1 = Remote, 2 = Hybrid, 3 = On Location)",
       y = "Work-Family Conflict Score") +
  theme_minimal()
```


```{r H-3: Work-family conflicts will be negatively correlated with Work-life balance}
# Perform Pearson correlation between Conflict Score and WLB Score
correlation_H3 <- cor(data$Conflict_score, data$WLB_score, use = "complete.obs", method = "pearson")

# Perform a test of significance for the correlation
cor_test_H3 <- cor.test(data$Conflict_score, data$WLB_score, method = "pearson")

# Display the results
correlation_H3
cor_test_H3

# Optional: Scatter plot to visualize the correlation
ggplot(data, aes(x = Conflict_score, y = WLB_score)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Correlation between Conflict Score and Work-Life Balance",
       x = "Conflict Score (WFC + FWC)",
       y = "Work-Life Balance Score") +
  theme_minimal()
```


